### 2018-04-12 algorithm.tex
new algo written up and coded. runs but probably buggy

\section{Full model}
Introduce a new variable for the factory-gate price of intermediate goods,
\begin{equation}\label{eq:varrho}
	\varrho_{njt} = \xi B_j T_{nj}^{-1/\theta} A_{njt}^{-1} w_{njt}^{\beta_j} \prod_{k=1}^J P_{nkt}^{\gamma_{kj}}.
\end{equation}
With this variable, we can write the price of the final good as
\begin{equation}\label{eq:price}
	P_{njt} = \left[
		\sum_{m=1}^N (\varrho_{mjt}/\kappa_{nmjt})^{-\theta}
		\right]^{-1/\theta}.
\end{equation}
This is a standard CES price index. We can write import shares as
\begin{equation}\label{eq:import_share}
	d_{nmjt} = \left[
	\frac
	{\varrho_{mjt}/\kappa_{nmjt}}
	{P_{njt}}
	\right]^{-\theta}.
\end{equation}
Wages can be expressed by inverting \eqref{eq:varrho},
\begin{equation}\label{eq:wage}
	w_{njt} = (\xi B_j)^{-1/\beta_j}
	\varrho_{njt}^{1/\beta_j}  
	T_{nj}^{1/(\beta_j\theta)}A_{njt}^{1/\beta_j} 
	\prod_{k=1}^J P_{nkt}^{-\gamma_{kj}/\beta_j}.
\end{equation}

Taking the market clearing condition,
\begin{equation}
	R_{mjt} = \sum_{n=1}^N
		d_{nmjt}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{nkt} - S_{nt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{nkt}
		\right],
\end{equation}
\[
	R_{mjt}\varrho_{mjt}^{\theta} = \sum_{n=1}^N
		(\kappa_{nmjt} P_{njt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{nkt} - S_{nt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{nkt}
		\right],
\]
\begin{equation}\label{eq:market_clearing}
	\varrho_{mjt} = R_{mjt}^{-1/\theta}
	\left\{
	\sum_{n=1}^N
		(\kappa_{nmjt} P_{njt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{nkt} - S_{nt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{nkt}
		\right]
	\right\}^{1/\theta}.
\end{equation}
Here both $R_{njt}$ and $\varrho_{njt}$ depend on equilibrium prices. But starting from a guess for $R$ (say, the autarky equilibrium), we can easily compute a shadow price $\varrho$ supporting that allocation. 
\begin{enumerate}
	\item Set $R_{njt}^{(0)}$ and $P_{njt}^{(0)}$ to the autarky equilibrium value or other suitable starting value.
	\item Given $R_{njt}^{(k)}$ and $P_{njt}^{(k)}$, solve for $\varrho_{njt}^{(k+1)}$ using equation \eqref{eq:market_clearing}.
	\item Solve for $w_{njt}^{(k+1)}$ and $P_{njt}^{(k+1)}$ using \eqref{eq:wage} and \eqref{eq:price}.
	\item Calculate $R_{njt}^{(k+1)} = w_{njt}^{(k+1)}L_{njt}/\beta_j$.
	\item Go back to Step 2 and repeat until convergence.
\end{enumerate}
Typically, this converges in 4--6 steps. This might even be a contraction mapping.

\subsection{Suitable starting values}
Suppose that trade is free, $\kappa_{nmjt}\equiv 1$. Then $P_{njt}\equiv P_{jt}$ and
\begin{equation*}
	\varrho_{mjt}^{(0)} = 
	P_{jt}
	R_{mjt}^{-1/\theta}
	\left\{
	\sum_{n=1}^N
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{nkt} - S_{nt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{nkt}
		\right]
	\right\}^{1/\theta}.
\end{equation*}
Let $R_{wjt}$ denote the world revenue in sector $j$.
\begin{equation*}
	\varrho_{mjt}^{(0)} = 
	P_{jt}
	R_{mjt}^{-1/\theta}
		\left[
				\sum_{k=1}^J
				(\alpha_{jt}\beta_k + \gamma_{jk}) R_{wkt}
		\right]^{1/\theta}.
\end{equation*}
FIXME: This only determines $\varrho_{mjt}/P_{jt}$, what determines $P_{jt}$ relative to other sectors?
\end{document}

>>>>

\section{Full model}
Introduce a new variable for the factory-gate price of intermediate goods,
\begin{equation}\label{eq:varrho}
	\varrho_{njt} = \xi B_j T_{nj}^{-1/\theta} A_{njt}^{-1} w_{njt}^{\beta_j} \prod_{k=1}^J P_{nkt}^{\gamma_{kj}}.
\end{equation}
With this variable, we can write the price of the final good as
\begin{equation}\label{eq:price}
	P_{mjt} = \left[
		\sum_{n=1}^N (\varrho_{njt}/\kappa_{mnjt})^{-\theta}
		\right]^{-1/\theta}.
\end{equation}
This is a standard CES price index. We can write import shares as
\begin{equation}\label{eq:import_share}
	d_{mnjt} = \left[
	\frac
	{\varrho_{njt}/\kappa_{mnjt}}
	{P_{mjt}}
	\right]^{-\theta}.
\end{equation}
Wages can be expressed by inverting \eqref{eq:varrho},
\begin{equation}\label{eq:wage}
	w_{njt} = (\xi B_j)^{-1/\beta_j}
	\varrho_{njt}^{1/\beta_j}  
	T_{nj}^{1/(\beta_j\theta)}A_{njt}^{1/\beta_j} 
	\prod_{k=1}^J P_{nkt}^{-\gamma_{kj}/\beta_j}.
\end{equation}

Taking the market clearing condition,
\begin{equation}
	R_{njt} = \sum_{m=1}^N
		d_{mnjt}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right],
\end{equation}
\[
	R_{njt}\varrho_{njt}^{\theta} = \sum_{m=1}^N
		(\kappa_{mnjt} P_{mjt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right],
\]
\begin{equation}\label{eq:market_clearing}
	\varrho_{njt} = R_{njt}^{-1/\theta}
	\left\{
	\sum_{m=1}^N
		(\kappa_{mnjt} P_{mjt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right]
	\right\}^{1/\theta}.
\end{equation}
Here both $R_{njt}$ and $\varrho_{njt}$ depend on equilibrium prices. But starting from a guess for $R$ (say, the autarky equilibrium), we can easily compute a shadow price $\varrho$ supporting that allocation. 
\begin{enumerate}
	\item Set $R_{njt}^{(0)}$ and $P_{njt}^{(0)}$ to the autarky equilibrium value or other suitable starting value.
	\item Given $R_{njt}^{(k)}$ and $P_{njt}^{(k)}$, solve for $\varrho_{njt}^{(k+1)}$ using equation \eqref{eq:market_clearing}.
	\item Solve for $w_{njt}^{(k+1)}$ and $P_{njt}^{(k+1)}$ using \eqref{eq:wage} and \eqref{eq:price}.
	\item Calculate $R_{njt}^{(k+1)} = w_{njt}^{(k+1)}L_{njt}/\beta_j$.
	\item Go back to Step 2 and repeat until convergence.
\end{enumerate}
Typically, this converges in 4--6 steps. This might even be a contraction mapping.

\begin{equation*}
	\varrho_{njt}/P_{njt} = 
	\left\{
	\sum_{m=1}^N
		(\kappa_{nmjt} P_{mjt}/P_{njt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt}/R_{njt} - S_{mt}/R_{njt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}/R_{njt}
		\right]
	\right\}^{1/\theta}.
\end{equation*}

\subsection{Suitable starting values}
Suppose that trade is free, $\kappa_{nmjt}\equiv 1$. Then $P_{njt}\equiv P_{jt}$ and
\begin{equation*}
	\varrho_{njt}^{(0)} = 
	P_{jt}
	R_{njt}^{-1/\theta}
	\left\{
	\sum_{m=1}^N
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right]
	\right\}^{1/\theta}.
\end{equation*}
Let $R_{wjt}$ denote the world revenue in sector $j$.
\begin{equation*}
	\varrho_{mjt}^{(0)} = 
	P_{jt}
	R_{mjt}^{-1/\theta}
		\left[
				\sum_{k=1}^J
				(\alpha_{jt}\beta_k + \gamma_{jk}) R_{wkt}
		\right]^{1/\theta}.
\end{equation*}
This only determines $\varrho_{mjt}/P_{jt}$, what determines $P_{jt}$ relative to other sectors?\begin{equation*}
	\varrho_{mjt}/P_{jt} = 
	\beta_j^{1/\theta}
	(w_{mjt}L_{mjt})^{-1/\theta}
		\left[
				\sum_{k=1}^J
				(\alpha_{jt} + \gamma_{jk}/\beta_k) 
				\sum_{n=1}^N w_{nkt}L_{nkt}
		\right]^{1/\theta}.
\end{equation*}
With no IO linkages,
\begin{equation*}
	\frac
	{\varrho_{mjt}}
	{P_{jt}} = 
	\alpha_{jt}^{1/\theta} 
	\left[
		\frac
		{w_{mjt}L_{mjt}}
		{\sum_{n=1}^N w_{nt}L_{nt}}
	\right]^{-1/\theta}.
\end{equation*}
Given equation \eqref{eq:price},
\[
1 = \sum_{m=1}^M (\varrho_{mjt}/P_{jt})^{-\theta}
\]
for all $j$. Substituting in,
\begin{equation*}
\alpha_{jt} = \frac
	{\sum_{m=1}^N
		{w_{mjt}L_{mjt}}}
		{\sum_{n=1}^N w_{nt}L_{nt}}
\end{equation*}
for all $j$. This pins down the scale of prices and wages in sector $j$.
\end{document}
